Supplementary Information -- Morphological convergence as on-line lexical analogy

### 2020-03-03 supplementary_information/si.tex
final commit? yeah right

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 25 14:32:38 2020
\begin{table}[ht]
\centering
\begin{tabular}{llll}
  \hline
verb class & irregular set & regular set & miscellaneous regular verbs \\ 
  \hline
BURNT & 6 & 42 & 1218 \\ 
  DROVE & 14 & 83 & 1218 \\ 
  KEPT & 12 & 31 & 1218 \\ 
  SANG & 8 & 13 & 1218 \\ 
  sum & 40 & 169 & 1218 \\ 
   \hline
\end{tabular}
\caption{Number of forms in each verb class, GCM training data} 
\end{table}
\subsection{Estimation}

@@

Note that the sets of exceptions and related forms of each rule can overlap, both respectively and with each other. As a consequence, the MGL rules apply to fewer forms than apparent from the table: 456 (instead of 617) in total.

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 25 14:32:38 2020
\begin{table}[ht]
\centering
\begin{tabular}{llrr}
  \hline
category & rule.type & related forms & exceptions \\ 
  \hline
BURNT & irregular &  24 &  41 \\ 
  DROVE & irregular &  21 & 114 \\ 
  KEPT & irregular &  10 &   9 \\ 
  SANG & irregular &  11 &  77 \\ 
  BURNT & regular &  38 &  21 \\ 
  DROVE & regular &  29 &  27 \\ 
  KEPT & regular &  67 &  41 \\ 
  SANG & regular &  47 &  38 \\ 
   \hline
\end{tabular}
\caption{Number of forms in each verb class, MGL training data} 
\label{mglsetsize}
\end{table}
\subsection{Estimation}

@@

As we see, however, the MGL also focuses the training set. The structural descriptions of the forms only cover a fraction of all verbs in the starting dictionary, organized by shared groups of segments. While disjunction increases the power of a theory (increasing the set of acceptable classes), disjunct classes frequently serve as input for phonological processes \citep{mielke2008emergence} . In the MGL, the disjunction emerges despite the lack of an initial specification: while one rule accounts for the specific behavior of a given form, multiple rules cover a given class. The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mgltable}.

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 25 14:32:38 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -0.11 & 0.16 & -0.67 & n.s. \\ 
  verb\_baseline\_average & 8.26 & 0.69 & 12.04 & *** \\ 
  co\_player\_lexical\_typicalityrandom & 0.04 & 0.17 & 0.25 & n.s. \\ 
  co\_player\_lexical\_typicalityreversed & 0.04 & 0.17 & 0.20 & n.s. \\ 
  participant\_prereg\_average & 7.13 & 0.56 & 12.67 & *** \\ 
  co\_player\_regularisation\_rate-40\% & -0.46 & 0.18 & -2.47 & * \\ 
  co\_player\_regularisation\_rate+40\% & 0.55 & 0.17 & 3.31 & *** \\ 
  verb\_baseline\_average:co\_player\_lexical\_typicalityrandom & -2.40 & 0.93 & -2.59 & ** \\ 
  verb\_baseline\_average:co\_player\_lexical\_typicalityreversed & -2.13 & 0.96 & -2.23 & * \\ 
   \hline
\end{tabular}
\caption{Experimental Data, Post-test Regression Model Summary. Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)} 
\label{regressiontable1}
\end{table}
Our regression model selection in this article proceeded as follows. Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}. Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function. However, our model tables (e.g., Tables 5 and 6 in the main text) include the significance levels generated by lme4 summaries.

We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors. We compare nested random effects structures as above, and retain any random slopes and intercepts which are supported by likelihood ratio tests \citep{baayen2008mixed}.

In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality}\footnote{In the SI, all p values rounded to 0.00 should be taken to indicate p $<$ 0.001.}.  Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}. We note that if {\em verb baseline mean} is {\em not} centered, collinearity is high (VIF=15.79) in this model. A non-centered model also yields a significant main effect for {\em lexical typicality}, with increased post-test regularization in the random and reversed conditions. However, we elect to take a more conservative approach by not including a potentially spurious main effect. Note that other than this main effect difference, models with and without centering yield results that are qualitatively indistinguishable; all other main effects and interactions are present in either case. The collinearity for the centered model (Table 5 in the main text) is acceptable, with a maximum VIF score of 2.9.

@@

We fit the \textsc{Celex-GCM} and the \textsc{Celex-MGL} on the baseline data to see how well they predict regularization responses by the baseline participants. The \textsc{Celex-MGL} outperforms the \textsc{Celex-GCM}, if slightly. Both models contribute independently to explaining variation in the data.

The concordance indices of the \textsc{Celex}-MGL (0.586) and the \textsc{Celex}-MGL (0.581) on the baseline data are similar, with some edge retained by the MGL, as in the case of \citep{albright2003rules}. Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 25 14:32:39 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -0.53 & 0.12 & -4.22 & *** \\ 
  rescaled\_CELEX\_mgl\_feature\_distance & 1.35 & 0.09 & 14.48 & *** \\ 
  rescaled\_CELEX\_gcm\_feature\_distance\_residual & 1.42 & 0.13 & 11.10 & *** \\ 
   \hline
\end{tabular}
\caption{Model 3: MGL scores and GCM residuals, baseline data} 
\label{resid1}
\end{table}
% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 25 14:32:40 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -1.23 & 0.16 & -7.68 & *** \\ 
  rescaled\_CELEX\_gcm\_feature\_distance & 2.39 & 0.15 & 16.06 & *** \\ 
  rescaled\_CELEX\_mgl\_feature\_distance\_residual & 1.01 & 0.09 & 11.15 & *** \\ 
   \hline
\end{tabular}
\caption{Model 4: GCM scores and MGL residuals, baseline data} 
\label{resid2}
\end{table}
Model 1 predicts the GCM scores from the MGL scores (M1: GCM regular score  $\sim$ MGL regular score). Model 2 predicts the MGL scores from the GCM scores (M2: MGL regular score  $\sim$ GCM regular score). In two additional models, we then use the residuals from each model (M1, M2) as an estimate of the variation present in one score that cannot be attributed to the other.  First, we use the residuals of Model 1 in another, mixed-effects logistic regression model along with the MGL scores to predict regular responses in the baseline task (M3: regular response  $\sim$ MGL score + GCM residual + (1 + MGL score + GCM residual$|$participant) + (1 $|$ verb)). We then residualize for the MGL. We use the residuals of Model 2 along with the GCM scores in a similar logistic regression model (M4: regular response  $\sim$ GCM score + MGL residual + (1 + GCM score + MGL residual $|$participant)).

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 25 14:32:40 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -2.95 & 0.29 & -10.01 & *** \\ 
  rescaled\_CELEX\_gcm\_feature\_distance & 4.37 & 0.42 & 10.28 & *** \\ 
  gcm\_score\_diff & 2.61 & 0.47 & 5.56 & *** \\ 
  participant\_prereg\_average & 5.29 & 0.60 & 8.79 & *** \\ 
   \hline
\end{tabular}
\caption{Model 5: Celex-GCM and individual score difference, post-test data} 
\label{indiv1}
\end{table}
% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 25 14:32:40 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -1.318 & 0.152 & -8.663 & *** \\ 
  rescaled\_CELEX\_mgl\_feature\_distance & 1.944 & 0.176 & 11.045 & *** \\ 
  mgl\_score\_diff & 0.304 & 0.139 & 2.181 & * \\ 
  participant\_prereg\_average & 7.203 & 0.591 & 12.190 & *** \\ 
   \hline
\end{tabular}
\caption{Model 6: Celex-MGL and individual score difference, post-test data} 
\label{indiv2}
\end{table}
We do this for both the GCM  (Model 5: regular response  $\sim$ \textsc{Celex} GCM score + extra GCM information + participant pre-test mean + (1 + \textsc{Celex} GCM + extra GCM information$|$participant) + (1$|$verb) and the MGL (Model 6: regular response  $\sim$ \textsc{Celex} MGL score + extra MGL information + participant pre-test mean + (1 + \textsc{Celex} MGL + extra MGL information$|$participant)+ (1$|$verb)).

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Feb 25 14:32:40 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -2.94 & 0.27 & -10.90 & *** \\ 
  r\_baseline\_mgl\_features & 1.60 & 0.17 & 9.36 & *** \\ 
  r\_individual\_gcm\_features & 2.79 & 0.38 & 7.37 & *** \\ 
  participant\_prereg\_average & 5.44 & 0.58 & 9.30 & *** \\ 
   \hline
\end{tabular}
\caption{Celex-MGL and Individual-GCM predict variation in the post-test} 
\label{comp3}
\end{table}
Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ 201.86) or individual-GCM ($\chi^2 = $ 81.82) as predictors. It is not significantly better if we include individual-MGL ($\chi^2 = $ 0.09) or {\sc Celex}-GCM ($\chi^2 = $ 2.03).

@@

\begin{figure}
\includegraphics{si-model_figure2}
\caption{Baseline and individual regularization weights: participants, GCM, MGL}
\label{bigfigure}
\end{figure}

>>>>

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Mar  3 13:14:57 2020
\begin{table}[ht]
\centering
\begin{tabular}{llll}
  \hline
verb class & irregular set & regular set & miscellaneous regular verbs \\ 
  \hline
BURNT & 6 & 42 & 1218 \\ 
  DROVE & 14 & 83 & 1218 \\ 
  KEPT & 12 & 31 & 1218 \\ 
  SANG & 8 & 13 & 1218 \\ 
  count of unique forms & 40 & 169 & 1218 \\ 
   \hline
\end{tabular}
\caption{Number of forms in each verb class, GCM training data} 
\label{gcmsetsize}
\end{table}
\subsection{Estimation}

@@

Note that the sets of exceptions and related forms of each rule can overlap, both respectively and with each other. As a consequence, the MGL rules apply to fewer forms than apparent from the table: 401 in total.

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Mar  3 13:14:57 2020
\begin{table}[ht]
\centering
\begin{tabular}{llrr}
  \hline
category & rule.type & related forms & exceptions \\ 
  \hline
BURNT & irregular &  24 &  41 \\ 
  DROVE & irregular &  21 & 114 \\ 
  KEPT & irregular &  10 &   9 \\ 
  SANG & irregular &  11 &  77 \\ 
  BURNT & regular &  38 &  21 \\ 
  DROVE & regular &  29 &  27 \\ 
  KEPT & regular &  67 &  41 \\ 
  SANG & regular &  47 &  38 \\ 
   \hline
\end{tabular}
\caption{Number of forms in each verb class, MGL training data} 
\label{mglsetsize}
\end{table}
\subsection{Estimation}

@@

As we see, however, the MGL also focuses the training set. The structural descriptions of the forms only cover a fraction of all verbs in the starting dictionary, organized by shared groups of segments. While disjunction increases the power of a theory (increasing the set of acceptable classes), disjunct classes frequently serve as input for phonological processes \citep{mielke2008emergence}. In the MGL, the disjunction emerges despite the lack of an initial specification: while one rule accounts for the specific behavior of a given form, multiple rules cover a given class. The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mglruletable}. Note that the regular rules themselves overlap: A regular rule that accounts for \emph{weep-weeped} can also account for \emph{burn-burned}. In total, there are 32 irregular rules and 12 regular rules posited by the MGL on the {\sc Celex} verbs.

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Mar  3 13:14:57 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \hline
category & irregular rule & regular rule & sum \\ 
  \hline
burnt &   4 &   2 &   6 \\ 
  drove &  13 &   3 &  16 \\ 
  kept &   5 &   5 &  10 \\ 
  sang &  10 &   4 &  14 \\ 
   \hline
\end{tabular}
\caption{Number of regular and irregular rules in each verb class, posited by the MGL for the training data (Note that the regular rules themselves overlap)} 
\label{mglruletable}
\end{table}

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Mar  3 13:14:57 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -0.11 & 0.16 & -0.67 & n.s. \\ 
  verb\_baseline\_average & 8.26 & 0.69 & 12.04 & *** \\ 
  co\_player\_lexical\_typicalityrandom & 0.04 & 0.17 & 0.25 & n.s. \\ 
  co\_player\_lexical\_typicalityreversed & 0.04 & 0.17 & 0.20 & n.s. \\ 
  participant\_prereg\_average & 7.13 & 0.56 & 12.67 & *** \\ 
  co\_player\_regularisation\_rate-40\% & -0.46 & 0.18 & -2.47 & * \\ 
  co\_player\_regularisation\_rate+40\% & 0.55 & 0.17 & 3.31 & *** \\ 
  verb\_baseline\_average:co\_player\_lexical\_typicalityrandom & -2.40 & 0.93 & -2.59 & ** \\ 
  verb\_baseline\_average:co\_player\_lexical\_typicalityreversed & -2.13 & 0.96 & -2.23 & * \\ 
   \hline
\end{tabular}
\caption{Experimental Data, Post-test Regression Model Summary. Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)} 
\label{regressiontable1}
\end{table}
Our regression model selection in this article proceeded as follows. Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}. Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function. However, our model tables (e.g., Tables 5 and 6 in the main text) include the Wald significance levels generated by lme4 summaries (* p$<$0.05; ** p$<$0.01; *** p$<$0.001).  
We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors. We compare nested random effects structures as above, and retain any random slopes and intercepts which are supported by likelihood ratio tests \citep{baayen2008mixed}.

In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality}.  Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}. We note that if {\em verb baseline mean} is {\em not} centered, collinearity is high (VIF=15.79) in this model. A non-centered model also yields a significant main effect for {\em lexical typicality}, with increased post-test regularization in the random and reversed conditions. However, we elect to take a more conservative approach by not including a potentially spurious main effect. Note that other than this main effect difference, models with and without centering yield results that are qualitatively indistinguishable; all other main effects and interactions are present in either case. The collinearity for the centered model (Table 5 in the main text) is acceptable, with a maximum VIF score of 2.9.

@@

We fit the \textsc{Celex-GCM} and the \textsc{Celex-MGL} on the baseline data to see how well they predict regularization responses by the baseline participants. The \textsc{Celex-MGL} outperforms the \textsc{Celex-GCM}, albeit only slightly. Both models contribute independently to explaining variation in the data.

The concordance indices of the \textsc{Celex}-MGL (0.586) and the \textsc{Celex}-MGL (0.581) on the baseline data are similar, with some edge retained by the MGL, as in the case of \cite{albright2003rules}. Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Mar  3 13:14:58 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -0.53 & 0.12 & -4.22 & *** \\ 
  rescaled\_CELEX\_mgl\_feature\_distance & 1.35 & 0.09 & 14.48 & *** \\ 
  rescaled\_CELEX\_gcm\_feature\_distance\_residual & 1.42 & 0.13 & 11.10 & *** \\ 
   \hline
\end{tabular}
\caption{Model 3: MGL scores and GCM residuals, baseline data} 
\label{resid1}
\end{table}
% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Mar  3 13:14:58 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -1.23 & 0.16 & -7.68 & *** \\ 
  rescaled\_CELEX\_gcm\_feature\_distance & 2.39 & 0.15 & 16.06 & *** \\ 
  rescaled\_CELEX\_mgl\_feature\_distance\_residual & 1.01 & 0.09 & 11.15 & *** \\ 
   \hline
\end{tabular}
\caption{Model 4: GCM scores and MGL residuals, baseline data} 
\label{resid2}
\end{table}
Model 1 predicts the GCM scores from the MGL scores (M1: GCM regular score  $\sim$ MGL regular score). Model 2 predicts the MGL scores from the GCM scores (M2: MGL regular score  $\sim$ GCM regular score). In two additional models, we then use the residuals from each model (M1, M2) as an estimate of the variation present in one score that cannot be attributed to the other.  First, we use the residuals of Model 1 in another, mixed-effects logistic regression model along with the MGL scores to predict regular responses in the baseline task (M3: regular response  $\sim$ MGL score + GCM residual + (1 + MGL score + GCM residual$|$participant) + (1 $|$ verb)). We then residualize for the MGL. We use the residuals of Model 2 along with the GCM scores in a similar logistic regression model (M4: regular response  $\sim$ GCM score + MGL residual + (1 + GCM score + MGL residual $|$participant)).

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Mar  3 13:14:58 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -2.95 & 0.29 & -10.01 & *** \\ 
  rescaled\_CELEX\_gcm\_feature\_distance & 4.37 & 0.42 & 10.28 & *** \\ 
  gcm\_score\_diff & 2.61 & 0.47 & 5.56 & *** \\ 
  participant\_prereg\_average & 5.29 & 0.60 & 8.79 & *** \\ 
   \hline
\end{tabular}
\caption{Model 5: Celex-GCM and individual score difference, post-test data} 
\label{indiv1}
\end{table}
% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Mar  3 13:14:58 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -1.32 & 0.15 & -8.66 & *** \\ 
  rescaled\_CELEX\_mgl\_feature\_distance & 1.94 & 0.18 & 11.04 & *** \\ 
  mgl\_score\_diff & 0.30 & 0.14 & 2.18 & * \\ 
  participant\_prereg\_average & 7.20 & 0.59 & 12.19 & *** \\ 
   \hline
\end{tabular}
\caption{Model 6: Celex-MGL and individual score difference, post-test data} 
\label{indiv2}
\end{table}
We do this for both the GCM  (Model 5: regular response  $\sim$ \textsc{Celex} GCM score + extra GCM information + participant pre-test mean + (1 + \textsc{Celex} GCM + extra GCM information$|$participant) + (1$|$verb) and the MGL (Model 6: regular response  $\sim$ \textsc{Celex} MGL score + extra MGL information + participant pre-test mean + (1 + \textsc{Celex} MGL + extra MGL information$|$participant)+ (1$|$verb)).

@@

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Tue Mar  3 13:14:59 2020
\begin{table}[ht]
\centering
\begin{tabular}{lrrrl}
  \hline
term & estimate & std.error & statistic & sig \\ 
  \hline
(Intercept) & -2.94 & 0.27 & -10.90 & *** \\ 
  rescaled\_CELEX\_mgl\_feature\_distance & 1.60 & 0.17 & 9.36 & *** \\ 
  rescaled\_individual\_gcm\_feature\_distance & 2.79 & 0.38 & 7.37 & *** \\ 
  participant\_prereg\_average & 5.44 & 0.58 & 9.30 & *** \\ 
   \hline
\end{tabular}
\caption{Celex-MGL and Individual-GCM predict variation in the post-test} 
\label{comp3}
\end{table}
Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ 201.86 ***) or individual-GCM ($\chi^2 = $ 81.82 ***) as predictors. It is not significantly better if we include individual-MGL ($\chi^2 = $ 0.09, p = 0.77) or {\sc Celex}-GCM ($\chi^2 = $ 2.03, p = 0.15).

@@

\begin{figure}
\centering
\includegraphics[width = \textwidth]{si_figure.pdf}
\caption{Baseline and individual regularization weights: participants, GCM, MGL}
\label{bigfigure}
\end{figure}
